\textbf{ Mapping stellar surfaces\\ I: Degeneracies in the rotational light curve problem }

### 2021-01-27 paper1/ms.tex
hogg comments

To date, most studies aimed at inferring stellar surface
properties from light curves follow one of two broad approaches. The
first is to model the stellar surface as a collection of one or more
discrete, circular, uniform contrast dark or bright spots on a uniform
intensity photosphere.
The advantage of this approach is that
the light curve can be computed efficiently and in some cases even
analytically \citep[e.g.,][]{Davenport2015,Morris2017,Morris2020b}.
%
The second approach is to discretize the surface at some resolution
and compute the emergent flux as a weighted sum of the visible pixel intensities.
This approach is more flexible, since it is not limited to surfaces
composed of distinct circular spots \citep[e.g.,][]{Harmon2000,Roettenbacher2017}.
%
Both approaches rely on an explicit \emph{forward model}, a prescription for
how to generate data given a set of parameters. In most cases, however, we are interested
in the inverse problem: constraining the parameters given data.
%
Unfortunately, the inverse problem is not only difficult---as it requires
a large number of forward model evaluations to find the parameter values
that are most consistent with the
data---but also formally \emph{ill-posed}.
%
While the mapping from a stellar surface to a light curve (the forward
problem) is unique, the mapping from a light curve to a surface
(the inverse problem) is not:
given any light curve, there exist an infinite number of surfaces that
could have generated it. These degeneracies are illustrated in
Figure~\ref{fig:degeneracies}, where six synthetic stellar surfaces
are shown (rows) at twelve different phases (columns), all at an
inclination $I=60^\circ$. Each surface
consists of a different number of dark spots on a brighter, heterogeneous
background.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/ylms.pdf}
        \oscaption{ylms}{%
            The real spherical harmonics in the polar frame up to
            $l = 5$, where dark colors correspond to negative intensity
            and bright colors to positive intensity.
            Rows correspond to the degree $l$ and columns to
            the order $m$. The set of all spherical harmonics forms a
            complete, orthogonal basis on the sphere.
            \label{fig:ylms}
        }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
        %
        \textbf{Single light curve at $I = 60^\circ$}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60d.pdf}
        \oscaption{nullspace_preimage}{%
            Decomposition of a surface map (left column) into its
            preimage (center) and null space (right) components
            for different surfaces, and their corresponding contributions
            to the rotational light curve. The preimage is the
            set of surfaces modes that map onto the light curve;
            the null space is the set of modes that do not.
            An inclination of $60^\circ$ is assumed in all plots.
            The vast majority of surface modes are in the null space
            of the light curve problem and therefore do not contribute
            to the observed flux.
            \label{fig:nullspace_preimage}
        }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        %
        \textbf{Single light curve at $I = 85^\circ$}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85d.pdf}
        \oscaption{nullspace_preimage}{%
            Same as Figure~\ref{fig:nullspace_preimage}, but for
            a stellar inclination of $85^\circ$. As the stellar rotation vector
            becomes perpendicular to the line of sight, it becomes
            more difficult to constrain latitudinal information.
            \label{fig:nullspace_preimage_85}
        }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
        %
        \textbf{Ensemble of light curves at many inclinations}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_d.pdf}
        \oscaption{nullspace_preimage_ensemble}{%
            Same as Figures~\ref{fig:nullspace_preimage} and \ref{fig:nullspace_preimage_85}, \
            but assuming we can measure the light curves of these stars from many
            different inclinations. In this limit, the information content of our
            data approaches $50\%$ of the spatial information about the surface.
            \label{fig:nullspace_preimage_ensemble}
        }
    \end{centering}
\end{figure}

>>>>

To date, most studies aimed at inferring stellar surface
properties from light curves follow one of two broad approaches. The
first is to model the stellar surface as a collection of one or more
discrete, circular, uniform contrast dark or bright spots on a uniform
intensity photosphere.
The advantage of this approach is that
the light curve can be computed efficiently and in some cases even
analytically \citep[e.g.,][]{Davenport2015,Morris2017,Morris2020b}.
%
The second approach is to discretize the surface at some resolution
and compute the emergent flux as a weighted sum of the visible pixel intensities.
This approach is more flexible, since it is not limited to surfaces
composed of distinct circular spots \citep[e.g.,][]{Harmon2000,Roettenbacher2017}.
%
Both approaches rely on an explicit \emph{forward model}, a prescription for
how to generate data given a set of parameters. In most cases, however, we are interested
in the inverse problem: constraining the parameters given data.
%
Unfortunately, the inverse problem is not only difficult---as it requires
a large number of forward model evaluations to find the parameter values
that are most consistent with the
data---but also formally \emph{ill-posed}.
%
While the mapping from a stellar surface to a light curve (the forward
problem) is unique, the mapping from a light curve to a surface
(the inverse problem) is not:
given any light curve, there exist an infinite number of surfaces that
could have generated it. These degeneracies are illustrated in
Figure~\ref{fig:degeneracies}, where six synthetic stellar surfaces
are shown (rows) at twelve different phases (columns), all at an
inclination $I=60^\circ$.%
\footnote{
    In this paper, we adopt the common convention where $I$ is the angle
    between the stellar spin axis and the line of sight,
    spanning $-90^\circ \leq I \leq 90^\circ$. An inclination $I=0^\circ$ therefore
    corresponds to a pole-on orientation.
}
Each surface
consists of a different number of dark spots on a brighter, heterogeneous
background.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/ylms.pdf}
        \oscaption{ylms}{%
            The real spherical harmonics in the polar frame
            ($\hat{\mathbf{x}}$ points to the right,
            $\hat{\mathbf{y}}$ points up, and $\hat{\mathbf{z}}$
            points out of the page)
            up to $l = 5$. Dark colors correspond to negative intensity
            and bright colors to positive intensity;
            rows correspond to the degree $l$ and columns to
            the order $m$. The set of all spherical harmonics forms a
            complete, orthogonal basis on the sphere.
            \label{fig:ylms}
        }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
        %
        \vspace{-4em}
        \includegraphics[width=1in]{figures/wireframe_60.pdf}
        \\[0.5em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_60d.pdf}
        \oscaption{nullspace_preimage}{%
            Decomposition of a surface map (left column) into its
            preimage (center) and null space (right) components
            for different surfaces, and their corresponding contributions
            to the rotational light curve. The preimage is the
            set of surfaces modes that map onto the light curve;
            the null space is the set of modes that do not.
            An inclination of $60^\circ$ is assumed in all plots.
            The vast majority of surface modes are in the null space
            of the light curve problem and therefore do not contribute
            to the observed flux.
            \label{fig:nullspace_preimage}
        }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        %
        \vspace{-4em}
        \includegraphics[width=1in]{figures/wireframe_85.pdf}
        \\[0.5em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_85d.pdf}
        \oscaption{nullspace_preimage}{%
            Same as Figure~\ref{fig:nullspace_preimage}, but for
            a stellar inclination of $85^\circ$. As the stellar rotation vector
            becomes perpendicular to the line of sight, it becomes
            more difficult to constrain latitudinal information.
            \label{fig:nullspace_preimage_85}
        }
    \end{centering}
\end{figure}

@@

\begin{figure}[p!]
    \begin{centering}
        %
        %
        \vspace{-4em}
        \includegraphics[width=1in]{figures/wireframe_30.pdf}
        \includegraphics[width=1in]{figures/wireframe_45.pdf}
        \includegraphics[width=1in]{figures/wireframe_60.pdf}
        \includegraphics[width=1in]{figures/wireframe_75.pdf}
        \\[0.5em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_a.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_b.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_c.pdf}
        \\[1em]
        \includegraphics[width=\linewidth]{figures/nullspace_preimage_d.pdf}
        \oscaption{nullspace_preimage_ensemble}{%
            Same as Figures~\ref{fig:nullspace_preimage} and \ref{fig:nullspace_preimage_85}, \
            but assuming we can measure the light curves of these stars from many
            different inclinations. In this limit, the information content of our
            data approaches $50\%$ of the spatial information about the surface.
            \label{fig:nullspace_preimage_ensemble}
        }
    \end{centering}
\end{figure}
